\documentclass[dvips, a4paper, 12pt,listof=totoc, oneside, parskip]{scrbook}

\usepackage{verbatim}
\usepackage[small]{caption}
\usepackage{graphicx,psfrag}
\usepackage{rotating}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage[dvips]{geometry,color}
\usepackage{xspace}
\usepackage[utf8]{inputenc}
\usepackage{wrapfig}
\usepackage{multirow}
\usepackage{placeins}

\title{Automapping rulesfile design description}
\date{Tiled v0.9}
\author{Stefan Beller}

\newcommand{\ind}{[\langle\textup{index}\rangle]}
\newcommand{\name}{\langle\textup{name}\rangle}
\begin{document}
\maketitle

\chapter*{abstract}
Automapping is an advanced tool to automatically search certain
combinations of tiles accross layers in a map and to replace
these parts by other combination. This allows the user to draw
structures with a minimum of time spent and the Automapping will be able
to generate a rather complex scenario, which would need lots more time if
manually crafted.

The first chapter of this document describes the Automapping rulemap
files in every detail.

The second chapter delivers examples which are cited by the the first part to
explain the usefulness of certain design decisions.


\chapter{Formal Description}
An automapping file consists of 4 major parts:

\begin{enumerate}
  \item The \emph{Definition of regions} describes which locations of the rulemap
  are actually used to create Automapping rules.

  \item The \emph{Definition of inputs} describes which kind of pattern the working map
  will be searched for.

  \item The \emph{Definition of outputs} describes how the working map is changed when
  an input pattern is found.

  \item The \emph{Map properties} are used to finetune the input pattern localization and the output of all rules within this rulesfile.
\end{enumerate}






\section{Defining the Regions}
There must be either a tilelayer called $\textup{regions}$ or there must be
the both tilelayers $\textup{regions}\_\textup{input}$ and
$\textup{regions}\_\textup{output}$

Using the regions layer, the region defined for input and output is
the same.

Using the different layers 'regions\_input' and 'regions\_output'
delivers the possibility to have different regions for the input section
and the output section.

The region layer(s) are only used to mark regions, where an automapping rule exists.
Therefore it does not matter which tiles are used in this layer, since these tiles are just used
to define a region.
So either use any tile or no tile at a coordinate to indicate if that coordinate belongs to a
rule or if it doesn't.

If multiple rules are defined in one rulemap file, the regions must not be
adjacent. That means there must be at least one tile of unused space in between
two rules. If the regions are adjacent (coherent) then both regions are interpreted as
one rule.

The use of different 'regions\_input' and 'regions\_output' is demonstrated
in example~\ref{example_tmw_grass_water_corners}.

\subsection{Multiple rules in one rulefile}

Of course multiple rules are possible in one rulemap.
If you want to have the rules applied in a certain sequence, you should
use multiple rulemaps and define the sequence within the rules.txt file.

As of now there also is a certain sequence within one rulemapfile.
Generally speaking the regions with small y value come first.
If there are regions at the same y value, then the x value is taken into account.
On orthogonal maps this ordering scheme is the same as for reading in most western countries.
(Left to right, top to down).

The order within one rulemap may be changed later,
once tiled is capable of utilizing multiple threads/processors.
So if you want to rely on a certain sequence, use different rulemaps and
order these in the rules.txt



\section{Definition of inputs}
Inputs are generally defined by tilelayers whichs name follows this scheme

$$\textup{input}[\textup{not}] [\langle index \rangle]  \_ \langle name \rangle$$

whereas the $[\textup{not}]$ and $[\langle index \rangle]$ are optional.

The $\langle name \rangle$ determines which layer on the working map is examined.
So for example the layer $\textup{input}\_\textup{Ground}$ will check the
layer called $\textup{Ground}$ in the working map for this rule.

Multiple layers having the same $\langle name \rangle$ and $\langle index \rangle$
is explicitly allowed and is intended. Having multiple layers of the same
$\langle name \rangle$ and $\langle index \rangle$, will allow you to define different possible tiles
per coordinate as input.

The $\langle index \rangle$ is used to create complete different input conditions.
All layers having the same $\langle index \rangle$ are taken into account for forming one
condition. Each of these conditions are checked individually.

\begin{enumerate}
  \item $\langle index \rangle$ must not contain an underscore.
  \item $\langle index \rangle$ must not start with 'not'
  \item $\langle index \rangle$ may be empty.
\end{enumerate}






\section{Definition of outputs}

Outputs are generally defined by tilelayers whichs name follows this scheme

$$\textup{output}[\langle index \rangle]  \_ \langle name \rangle$$

which is very similar to the input section. At first there must be the word $\textup{output}$.
Then optionally an $\textup{index}$ may occur. After the first underscore there will be the
$\textup{name}$ of the target layer.

All layers of the same index are treated as one possible output.
So the intention of indexes in the outputs of rules is only used for random
output.

The indexes in the output section have nothing to do with the indexes in the input section,
they are independant. In the output section they are used for randomness; in the input section
they are used to define multiple possible layers as input.

So when there are multiple indexes within one rule, the output will be choosen
fairly (uniformly distributed) accross all indexes. So a dice will be rolled and one index is picked.
All of the output layers carrying this index will be put out into the working map then.

Note that the output is not being checked for overlapping itself. This can be archieved by
setting the map property \emph{NoOverlappingRules} to true.




\section{Map properties}

There are 3 different map properties, which can be used to add additional
information to a rulefile:
\begin{enumerate}
  \item \emph{NoOverlappingRules}
  \item \emph{DeleteTiles}
  \item \emph{AutomappingRadius}
\end{enumerate}
These properties are map wide, meaning it applies to all rules which are part of the rulemap.
If you need rules with different properties, you can use multiple rulemaps.


\subsection{NoOverlappingRules}
A rule may not overlap to itself. It is used in example~\ref{alternatingWall}

\subsection{DeleteTiles}
This deletes all tiles in all possibly changed layers before the automapping takes place.
This is useful when using live Automapping in a map with layers, which only carry
a few tiles.

\subsection{AutomappingRadius}
Only interesting in live Automapping.

In live Automapping only a small part of the map is taken into account for Automapping to speed
up the process.
Since some rules rely on other rules to be executed first, these rules maybe should be
applied in a bigger region.

\chapter{Examples}

\input{Example/TheManaWorld/themanaworld}

\input{Example/LoneCoder/alternatingwalls}

\input{Example/AbstractInput/AbstractInput}

\chapter{Hints}

dedicated set layer or trying to parse your layer

\end{document}
